### 2018-05-29 abstract.tex
Mearge revision from Prof. Xue

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  have been used to effectively exploit the capability of
  on-chip lastly-recent-used (LRU) caches,
  which are popular in traditional high-performance
  processors used in high-end servers and supercomputers.
  However, the market has recently witnessed a
  new diversity in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to the development of
  high-performance GEMM in a multi-threaded context.
  As several threads try to load data into a shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in the GEMM routines,
  by partitioning a shared cache into physically disjoint
  sets and assigning different sets to different threads.
  We have implemented SCP in the OpenBLAS library and
  evaluated it on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared pseudo-random L2 caches (per four-core cluster).
  Our evaluation shows that SCP has effectively reduced the
  conflict misses in both L1 and L2 caches,
  resulting in an improvement of the GEMM performance by $2.75\%$ -- $6.91\%$.
\end{abstract}

>>>>

\begin{abstract}
  GEneral Matrix Multiply (GEMM) is the most fundamental
  computational kernel routine in the BLAS library.
  To achieve high performance, in-memory data must be
  prefetched into fast on-chip caches before they are used.
  Two techniques, software prefetching and data packing,
  have been used to effectively exploit the capability of
  on-chip lastly-recent-used (LRU) caches,
  which are popular in traditional high-performance
  processors used in high-end servers and supercomputers.
  However, the market has recently witnessed a
  new diversity in processor design, resulting in high-performance processors
  equipped with shared caches with non-LRU replacement policies.
  This poses a challenge to the development of
  high-performance GEMM in a multi-threaded context.
  As several threads try to load data into a shared cache simultaneously,
  inter-thread cache conflicts will increase significantly.
  We present a Shared Cache Partitioning (SCP) method to
  eliminate inter-thread cache conflicts in the GEMM routines,
  by partitioning a shared cache into physically disjoint
  sets and assigning different sets to different threads.
  We have implemented SCP in the OpenBLAS library and
  evaluated it on Phytium 2000+,
  a 64-core AArch64 processor with private LRU L1 caches
  and shared pseudo-random L2 caches (per four-core cluster).
  Our evaluation shows that SCP has effectively reduced the
  conflict misses in both L1 and L2 caches in a highly-optimized GEMM implementation,
  resulting in an improvement of its performance by $2.75\%$ -- $6.91\%$.
\end{abstract}

### 2018-05-29 background.tex
Formatting adjust and some minor update
- make the figures larger
- add a new table for software environment in evaluation
- add two horizontal lines on the breakdown figure in the privatization section.
- textual updates highlighted in red color
- add funding information
- add PLASMA references
Mearge revision from Prof. Xue

A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
where $c$ is the cache capacity, $l$ is the cache line size,
$n$ is the number of ways-of-associativity,
and $nt$ indicates the number of cores sharing the cache.
The whole cache memory is divided into $ns=c/l/n$ sets,
identified by indices in $\mathcal{S} = [0, ns)$,
with each set containing $n$ cache lines.
Generally, $ns$ is a power-of-two.
An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
mapping the addresses in the address space $\mathcal{A}$
(either virtual or physical) to the cache sets in
$\mathcal{S}$.
The data at $addr \in \mathcal{A}$ must be fetched
into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
A classic implementation of the indexing function is shown below,
%% $\varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)$,
with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively:
\begin{equation*}
  \varphi(addr) = (addr \gg log_2(l)) ~\&~ (ns-1)
  \label{eq:phi}
\end{equation*}
which represents an extraction of the $log_2(ns)$ bits starting from the $log_2(l)$-th
position in $addr$'s binary form.
It is easy to prove that any continuous memory block
that is no larger than $c/n$
cannot conflict with itself.
We therefore define a quantity, \emph{way-capacity}, denoted as $wc=c/n$,
to represent this size bound for conflict-free memory blocks.

@@

Throughout this paper, our example platform will be
a 4-core processor with
$NL=2$ levels of caches.
Figure~\ref{fig:hierarchy} shows its memory hierarchy 
and Table~\ref{tab:cluster} lists its
architectural parameters. 
Each core has 32 128-bit (16B) vector registers,
each capable of storing two double-precision floating-point numbers.
For convenience, a processor core at level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with a programmable replacement policy.

\begin{figure}
  \centering
  \includegraphics[width=.45\textwidth]{figures/cluster-new}
  \caption{Memory hierarchy of the 4-core processor
    (nodes labeled with $(layer,index)$)}
  \label{fig:hierarchy}
\end{figure}

\begin{table}
  \centering
  \caption{Architectural parameters of the 4-core processor}
  %% (see Section~\ref{subsec:hierarchy} for meanings of symbols in the header row)
  \label{tab:cluster}
  \begin{tabular}{lccccccl}
    \toprule
    Type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & Policy \\
    \midrule
    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & Programmable \\
    L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & Pseudo-Random \\
    \bottomrule
  \end{tabular}
\end{table}

@@

In this blocked algorithm, the $jj$, $kk$ and $ii$ loops
along the $N$, $K$ and $M$ matrix dimensions are
tiled by sizes $N_c$, $K_c$ and $M_c$ at layers 1, 2 and 3, respectively.
At layers 4 and 5, the $j$ and $i$ loops along the $N$ and $M$ dimensions
are further tiled by sizes $N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ sub-matrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices at each level fit into a certain level in the memory hierarchy, with the following
constraints:
\begin{eqnarray}
  es (M_r + N_r + M_r N_r) & \le & c_{0} \label{eq:constraints.reg}\\
  nt_{1} \cdot es (N_r K_c + 2 M_r K_c) & \le & c_{1} \label{eq:constraints.l1}\\
  nt_{2} \cdot es (M_c K_c + 2 N_r K_c) & \le & c_{2} \label{eq:constraints.l2}\\
  es (N_c K_c + nt_{3} \cdot M_c K_c)   & \le & c_{3} \label{eq:constraints.l3}
\end{eqnarray}
where $es$ denotes the size of a matrix element,
e.g., 8B for a double-precision floating-point number
and $c_l$ denotes the capacity of the register file or cache at level $l$.
By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can
fit into the registers (the pseudo level-0 cache).
By (\ref{eq:constraints.l1}), $K_c$ is so constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
at layer 5 fit into the L1 cache.
By (\ref{eq:constraints.l2}), $M_c$ is so constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
at layer 4 fit into the L2 cache.
Finally, by (\ref{eq:constraints.l3}), $N_c$ is so constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache (if it exists).
Larger matrices at higher layers 1 -- 3 always reside in main memory.

@@

As a result, there are multiple $A_2$ instances (one per thread)
and a single $B_2$ instance (shared by all threads).
To understand how the $ii$ loop at layer 3 is parallelized on the 4-core processor, 
Figure~\ref{fig:workload} shows the data accessing patterns
for a single thread $T_1$.
While all the color shaded submatrices are accessed by $T_1$,
only the submatrices highlighted by the z-curves
(a subset of color shaded ones) are packed by $T_1$.
In other words, each thread packs its own $A_2$ instance but all the threads pack collaboratively the unique $B_2$ instance.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

\begin{figure}[t]
  \centering
  \includegraphics[width=.4\textwidth]{figures/workload}
  \caption{GEMM parallelization with 4 threads and data accessing patterns of thread $T_1$}
  \label{fig:workload}
\end{figure}

@@

In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in the 
L2 cache during all the $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar in that it is expected to
reside in the L1 cache during all the
$M_c/M_r$ iterations at layer~5.
In general, even with carefully selected tile factors,
some matrix elements can still be evicted.
For example, some data structures other than the packed matrices
or program code (in case of a unified cache) may be fetched into the
cache set where these matrix elements reside in.
Such cache conflicts are known as
\emph{intra-thread cache conflicts}
because they occur within the execution of a single thread.
To deal with intra-thread cache conflicts,
the GEBP kernel repeatedly prefetches data from $A_4$ and $B_4$ to the L1 cache
during every iteration at layer 6 before it actually reads them just in case that
the data may have been evicted accidentally from the L1 cache.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefetching strategy
can reduce the penalty caused by intra-thread cacheconflicts.

In contrary to intra-thread cache conflicts,
there exists another kind of cache conflicts, called
\emph{inter-thread cache conflicts}, on architectures with shared caches.
Suppose two working threads, $T_0$ and $T_1$, are running independently
on the 4-core processor in Figure.~\ref{fig:hierarchy}
to accomplish a GEMM operation.
At time $t_s$, $T_0$ prefetches some data at address $addr_0$
to the cache set $\varphi_2(addr_0)$ in a shared L2 cache.
The data will be actually read at time $t_e$.
Then at some time $t_m$ ($t_s < t_m < t_e$),
$T_1$ prefetches data at address $addr_1$ to the L2 cache.
If $\varphi_2(addr_1)=\varphi_2(addr_0)$, then the
data at $addr_1$ and $addr_0$ will be fetched into the same cache set.
As the shared L2 cache has a non-LRU policy,
the data at $addr_0$ may suffer an eviction before it is actually read,
i.e., a conflict miss occurs.

>>>>

A set-associative cache can be defined by a 4-tuple $(c, l, n, nt)$,
where $c$ is the cache capacity, $l$ is the cache line size,
$n$ is the number of ways-of-associativity,
and $nt$ indicates the number of cores sharing the cache.
The whole cache memory is divided into $ns=c/l/n$ sets,
identified by indices in $\mathcal{S} = [0, ns)$,
with each set containing $n$ cache lines.
Generally, $ns$ is a power-of-two.
An indexing function $\varphi: \mathcal{A} \mapsto \mathcal{S}$ is responsible for
mapping the addresses in the address space $\mathcal{A}$
(either virtual or physical) to the cache sets in
$\mathcal{S}$.
The data at $addr \in \mathcal{A}$ must be fetched
into the cache set $\varphi(addr) \in \mathcal{S}$ if requested.
A classic implementation of the indexing function is shown below,
%% $\varphi(addr) = (addr \gg \log_2(l)) ~\&~ (ns-1)$,
with $\gg$ and $\&$ representing logical-shift-right and bitwise-and operations, respectively:
\begin{equation*}
  \varphi(addr) = (addr \gg \log_2(l)) ~\&~ (ns-1)
  \label{eq:phi}
\end{equation*}
which represents an extraction of the $\log_2(ns)$ bits starting from the $\log_2(l)$-th
position in $addr$'s binary form.
It is easy to prove that any continuous memory block
that is no larger than $c/n$
cannot conflict with itself.
We therefore define a quantity, \emph{way-capacity}, denoted as $wc=c/n$,
to represent this size bound for conflict-free memory blocks.

@@

Throughout this paper, our example platform will be
a four-core processor with
$NL=2$ levels of caches.
Figure~\ref{fig:hierarchy} shows its memory hierarchy 
and Table~\ref{tab:cluster} lists its
architectural parameters. 
Each core has 32 128-bit (16B) vector registers,
each capable of storing two double-precision floating-point numbers.
For convenience, a processor core at level 0
is represented by its registers, which can be viewed as a special
level-0 cache $L0 = (512B, 16B, 32, 1)$ with a programmable replacement policy.

\begin{figure}
  \centering
  \includegraphics[width=.65\textwidth]{figures/cluster-new}
  \caption{Memory hierarchy of the four-core processor
    (nodes labeled with $(layer,index)$)}
  \label{fig:hierarchy}
\end{figure}

\begin{table}
  \centering
  \caption{Architectural parameters of the four-core processor}
  %% (see Section~\ref{subsec:hierarchy} for meanings of symbols in the header row)
  \label{tab:cluster}
  \begin{tabular}{lccccccl}
    \toprule
    Type & $c$ & $n$ & $l$ & $nt$ & $ns$ & $wc$ & Policy \\
    \midrule
    L0 Registers  & 512B & 32 & 16B & 1 & 1 & 16B & Programmable \\
    L1 Data    & 32KB & 2  & 64B & 1 & 256 & 16KB & LRU \\
    L2 Unified & 2MB  & 16 & 64B & 4 & 2048 & 128KB & Pseudo-Random \\
    \bottomrule
  \end{tabular}
\end{table}

@@

In this blocked algorithm, the $jj$, $kk$ and $ii$ loops
along the $N$, $K$ and $M$ matrix dimensions are
tiled by sizes $N_c$, $K_c$ and $M_c$ at layers 1, 2 and 3, respectively.
At layers 4 and 5, the $j$ and $i$ loops along the $N$ and $M$ dimensions
are further tiled by sizes $N_r$ and $M_r$, respectively.
As a result, the innermost loop at layer 6 goes over the $K$
dimension for a total of $K_c$ times,
with each iteration performing a rank-1 update on
the $M_r \times N_r$ sub-matrix of $C$, as shown at layer 7.
In OpenBLAS, matrix $C$ is scaled by $\beta$ before the loops,
so the multiply-accumulate operation at layer 7 does not have to scale $C$ again.
Tiling factors $N_c$, $K_c$, $M_c$, $N_r$ and $M_r$ are carefully selected so that
matrices at each level fit into a certain level in the memory hierarchy, with the following
constraints:
\begin{eqnarray}
  es (M_r + N_r + M_r N_r) & \le & c_{0} \label{eq:constraints.reg}\\
  nt_{1} \cdot es (N_r K_c + 2 M_r K_c) & \le & c_{1} \label{eq:constraints.l1}\\
  nt_{2} \cdot es (M_c K_c + 2 N_r K_c) & \le & c_{2} \label{eq:constraints.l2}\\
  es (N_c K_c + nt_{3} \cdot M_c K_c)   & \le & c_{3} \label{eq:constraints.l3}
\end{eqnarray}
where $es$ denotes the size of a matrix element,
e.g., 8B for a double-precision floating-point number
and $c_l$ ($0\leqslant l \leqslant 3$) denotes the capacity of the register file or cache at level $l$.
By (\ref{eq:constraints.reg}), $M_r$ and $N_r$ are so
constrained that $M_r$ elements from $A$,
$N_r$ elements from $B$ and $M_r \times N_r$ elements from $C$ can
fit into the registers (the pseudo level-0 cache).
By (\ref{eq:constraints.l1}), $K_c$ is so constrained that $B_4$ ($N_r \times K_c$),
$A_4$ ($M_r \times K_c$), and $A_4$'s counterpart in the next iteration
at layer 5 fit into the L1 cache.
By (\ref{eq:constraints.l2}), $M_c$ is so constrained that $A_3$ ($M_c \times K_c$),
$B_3$ ($N_r \times K_c$), and $B_3$'s counterpart in the next iteration
at layer 4 fit into the L2 cache.
Finally, by (\ref{eq:constraints.l3}), $N_c$ is so constrained that 
$B_2$ ($K_c \times N_c$) and $A_2$ ($M_c \times K_c$) fit into the L3 cache (if it exists).
Larger matrices at higher layers 1 -- 3 always reside in main memory.

@@

As a result, there are multiple $A_2$ instances (one per thread)
and a single $B_2$ instance (shared by all threads).
To understand how the $ii$ loop at layer 3 is parallelized on the four-core processor, 
Figure~\ref{fig:workload} shows the data accessing patterns
for a single thread $T_1$.
While all the color shaded submatrices are accessed by $T_1$,
only the submatrices highlighted by the z-curves
(a subset of color shaded ones) are packed by $T_1$.
In other words, each thread packs its own $A_2$ instance but all the threads pack collaboratively the unique $B_2$ instance.
%% We will use the OpenBLAS GEMM implementation in Figure.~\ref{fig:gemm}
%% in discussion and evaluation throughout the paper.

\begin{figure}[t]
  \centering
  \includegraphics[width=.5\textwidth]{figures/workload}
  \caption{GEMM parallelization with four threads and data accessing patterns of thread $T_1$}
  \label{fig:workload}
\end{figure}

@@

In Figure.~\ref{fig:gemm}, the packed matrix $A_2$
will be read $N_c/N_r$ times within a single GEBP execution,
and is expected to reside in the 
L2 cache during all the $N_c/N_r$ iterations at layer 4.
The packed matrix $B_4$ is similar in that it is expected to
reside in the L1 cache during all the
$M_c/M_r$ iterations at layer~5.
In general, even with carefully selected tile factors,
some matrix elements can still be evicted.
For example, some data structures other than the packed matrices
or program code (in case of a unified cache) may be fetched into the
cache set where these matrix elements reside in.
Such cache conflicts are known as
\emph{intra-thread cache conflicts}
because they occur within the execution of a single thread.
To deal with intra-thread cache conflicts,
the GEBP kernel repeatedly prefetches data from $A_4$ and $B_4$ to the L1 cache
during every iteration at layer 6 before it actually reads them just in case that
the data may have been evicted accidentally from the L1 cache.
As GEBP accesses the matrix data much more frequently than other
data structures, this repeated prefetching strategy
can reduce the penalty caused by intra-thread cache conflicts.

In contrary to intra-thread cache conflicts,
there exists another kind of cache conflicts, called
\emph{inter-thread cache conflicts}, on architectures with shared caches.
Suppose two working threads, $T_0$ and $T_1$, are running independently
on the four-core processor in Figure.~\ref{fig:hierarchy}
to accomplish a GEMM operation.
At time $t_s$, $T_0$ prefetches some data at address $addr_0$
to the cache set $\varphi_2(addr_0)$ in the shared L2 cache.
The data will be actually read at time $t_e$.
Then at some time $t_m$ ($t_s < t_m < t_e$),
$T_1$ prefetches data at address $addr_1$ to the L2 cache.
If $\varphi_2(addr_1)=\varphi_2(addr_0)$, then the
data at $addr_1$ and $addr_0$ will be fetched into the same cache set.
As the shared L2 cache has a non-LRU policy,
the data at $addr_0$ may suffer an eviction before it is actually read,
i.e., a conflict miss occurs.

### 2018-05-29 evaluation.tex
Mearge revision from Prof. Xue

We have implemented SCP in the OpenBLAS~\cite{openblas} library
%% (version 0.3.0-dev)
and evaluated it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing 4 cores.
The structure of a cluster is almost the same as
the 4-core processor in Figure~\ref{fig:hierarchy}
except that the L2 caches of all 16 clusters connect to the main memory
with hardware coherence.
\textcolor{red}{
Table~\ref{tab:cluster} lists the
architectural parameters for the cores and caches
in this processor
and Table~\ref{tab:software} shows the software environment used in out evaluation.}
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers~\cite{top500},
relies on the DGEMM variant.

@@

Figure~\ref{fig:privb} shows the results for partial
privatization with $NT_C=4$ and $NC=4$.
The results for the other 
configurations are similar.
For comparison purposes,
SCP with and without privatization,
denoted as SCP-P and SCP, respectively,
are both given in Figure~\ref{fig:privb}.
Figure~\ref{fig:privb.papi} shows the L1 and L2
cache miss rates, which are 
slightly lower with privatization (SCP-P) than without
(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower L1 and L2 cache miss rates,
privatization (SCP-P) suffers a performance loss
(relative to SCP) due to 
the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} compares the
three types of major overheads in percentage terms 
over the total execution time, including
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
\textcolor{red}{
The solid and dashed horizontal lines represent
the average value of total overhead across matrix sizes
under SCP and SCP-P, respectively.}
While behaving similarly as SCP
in (1) and (3), SCP-P is worse than SCP in (2).
To understand this further,
Table~\ref{tab:breakdown} gives a breakdown of each's
execution time in terms of (1) -- (3) and the
computation time of GEBP. As
the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
the total overhead has thus increased
from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).

>>>>

We have implemented SCP in the OpenBLAS~\cite{openblas} library
%% (version 0.3.0-dev)
and evaluated it on a Phytium 2000+ processor.
The Phytium 2000+ processor is an emerging high-performance
64-core processor based on ARM's AArch64 architecture.
The 64 cores are organized into 16 clusters with each
cluster containing four cores.
The structure of a cluster is almost the same as
the four-core processor in Figure~\ref{fig:hierarchy}
except that the L2 caches of all 16 clusters connect to the main memory
with hardware coherence.
Table~\ref{tab:cluster} lists the
architectural parameters for the cores and caches
in this processor
and Table~\ref{tab:software} shows the software environment used.
We restrict our evaluation to DGEMM,
as in prior work~\cite{blispar,augem,poetmicro}, for two reasons.
First, the basic idea behind SCP applies to other
variants of GEMM such as SGEMM, CGEMM and ZGEMM.
Second, the LINPACK benchmark, which is used to build the
TOP500 list of world's most powerful supercomputers~\cite{top500},
relies on the DGEMM variant.

@@

Figure~\ref{fig:privb} shows the results for partial
privatization with $NT_C=4$ and $NC=4$.
The results for the other 
configurations are similar.
For comparison purposes,
SCP with and without privatization,
denoted as SCP-P and SCP, respectively,
are both given in Figure~\ref{fig:privb}.
Figure~\ref{fig:privb.papi} shows the L1 and L2
cache miss rates, which are 
slightly lower with privatization (SCP-P) than without
(SCP). Figure~\ref{fig:privb.ate} shows the average thread efficiency.
Despite its lower L1 and L2 cache miss rates,
privatization (SCP-P) suffers a performance loss
(relative to SCP) due to 
the extra packing overhead introduced by privatization.
Figure~\ref{fig:privb.breakdown} compares the
three types of major overheads in percentage terms 
over the total execution time, including
(1) packing of $A_2$, (2) packing of $B_2$,
and (3) synchronization.
The solid and dashed horizontal lines represent
the average values of the total overheads
across the matrix sizes
under SCP and SCP-P, respectively.
While behaving similarly as SCP
in (1) and (3), SCP-P is worse than SCP in (2).
To understand this further,
Table~\ref{tab:breakdown} gives a breakdown of each's
execution time in terms of (1) -- (3) and the
computation time of GEBP. As
the overhead for packing $B_2$
is multiplied by a factor of $NT_C=4$,
the total overhead has thus increased
from $2.35\%$ (under SCP) to $4.72\%$ (under SCP-P).

### 2018-05-29 intro.tex
Mearge revision from Prof. Xue

The main contributions of this paper are as follows:
\begin{itemize}
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on the GEMM performance.
\item We propose SCP, a method for solving the inter-thread cache conflicts
  problem on architectures with shared non-LRU caches.
\item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and shared L2 caches with a pseudo-random
  replacement policy.
  Our evaluation shows that SCP can improve the GEMM performance consistently
  under various parallelization configurations, by $2.75\%$--$6.91\%$.
\end{itemize}

>>>>

The main contributions of this paper are as follows:
\begin{itemize}
\item We propose SCP, a method for reducing 
	the inter-thread cache conflicts
  for GEMM on architectures with shared non-LRU caches.
\item We present a quantitative analysis of the negative effect of inter-thread cache
  conflicts on the GEMM performance
  on architectures with shared non-LRU caches.
\item We have implemented SCP in the OpenBLAS library and evaluated it on Phytium 2000+,
  an emerging high-performance processor based on ARM's AArch64 architecture.
  Phytium 2000+ has 64 cores, private LRU L1 caches
  and shared L2 caches with a pseudo-random
  replacement policy. By reducing the inter-thread
  cache conflicts effectively,
  SCP can improve the 
  performance of a highly-optimized
  GEMM implementation  consistently
  under various parallelization configurations, by 
  $2.75\%$ -- $6.91\%$.
\end{itemize}

### 2018-05-29 related.tex
Mearge revision from Prof. Xue

>>>>

As GEMM is highly optimized and parallelized,
other higher-level dense linear algebra operations,
such as Cholesky, LU and QR factorization, can
enjoy its performance benefits by building themselves
on top of GEMM.
In other words, this programming pattern in dense linear algebra software
harness a fork-and-join style parallelism.
To reduce synchronization overhead,
the PLASMA project~\cite{plasma2009,plasma2010,plasma2017}
makes use of a task-driven parallelization paradigm,
in which dense linear algorithms are formulated in
terms of 
directed-acyclic-graphs (DAGs) of fine-grained tasks.
The tasks are dynamically scheduled based on their
data dependencies (represented by the graph edges).
The basic idea behind SCP also applies to such
DAG-based parallelization.
Input matrices of different tasks can be packed into
different sets of a shared cache. As a result, 
the tasks can be scheduled based on the cache sets 
where the input matrices reside in, so that
the tasks whose input matrices reside 
in the same cache set
are scheduled to the same core.
